---
title: bedbaser
author:
- name: J Wokaty
  affiliation: CUNY School of Public Health
  email: jennifer.wokaty@sph.cuny.edu
package: bedbaser
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{bedbaser}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

# Introduction

```{r library, include=FALSE}
library(bedbaser)
library(rtracklayer)
library(R.utils)
library(liftOver)
```

bedbaser is an R api for [BEDbase](https://bedbase.org) that provides access
to the [BEDbase API](https://api.bedbase.org) and includes convenience
functions, such for creating GRanges and GRangesList objects.

# Install bedbaser

Install bedbaser using `r CRANpkg("BiocManager")`.

```{r install, eval=FALSE}
if (!"BiocManager" %in% rownames(installed.packages())) {
    install.packages("BiocManager")
}
BiocManager::install("bedbaser", dependencies = TRUE)
```

# Create a bedbaser api

bedbaser uses the `r Biocpkg("AnVIL")` Service class to create a `BEDbase`
api, which allows access to the API directly.

```{r bedbase}
api <- bedbaser()
show(api)
```

End points can be accessed with `$` and their results can be retrieved
with `r CRANpkg("httr")`.

```{r count_bed_record_bed_count_get, eval=FALSE}
library(httr)
rsp <- api$get_example_bed_record_v1_bed_example_get()
ex_bed <- content(rsp)
```

bedbaser includes convenience functions prefixed with *bb_* to facilitate
finding BED files, exploring their metadata, and download files.

## Set the cache

Set the cache path with the environment variable `BEDBASER_CACHE`; otherwise,
bedbaser will choose the default cache location.

```{r BEDBASER_CACHE, eval=FALSE}
Sys.setenv("BEDBASER_CACHE" = tempdir())
```

# Find a BED file or BEDset

Use `bb_list_beds()` and `bb_list_bedsets()` to browse available resources in
BEDbase. Both functions display the record identifiers and names of BED files 
and BEDsets. A record identifier can be used to access a specific resource.

```{r bb_list_beds}
bb_list_beds(api)

bb_list_bedsets(api)
```

# Examine metadata

Use `bb_metadata()` to learn more about a BED or BEDset associated with
a record identifier.

```{r bb_metadata}
ex_bed <- bb_example(api, "bed")
md <- bb_metadata(api, ex_bed$id)
head(md)
```

# Show BED files in BEDset

Use `bb_beds_in_bedset()` to display the record identifiers of BEDs in a
BEDset. 

```{r bb_beds_in_bedset}
bb_beds_in_bedset(api, "excluderanges")
```

# Search for a BED file by keyword

Search for BED files by keywords. `bb_bed_text_search()` returns all BED files
scored against a query.

```{r bb_search}
bb_bed_text_search(api, "cancer", limit = 10)
```

# Import a BED into a GRanges object

Create a GRanges object with a BED record identifier with `bb_to_granges`, which
downloads and imports a BED file using `rtracklayer`. `bb_to_granges` also
attempts to pass the genome to `rtracklayer`

```{r bb_to_granges, message=FALSE}
ex_bed <- bb_example(api, "bed")
head(ex_bed)
# Allow bedbaser to assign column names and types
bb_to_granges(api, ex_bed$id, quietly = FALSE)
```

For BEDX+Y formats, a named list with column types may be passed through
`extra_cols` if the column name and type are known. Otherwise, `bb_to_granges`
guesses the column types and assigns column names.

```{r bb_to_granges_manual, eval=FALSE}
# Manually assign column name and type using `extra_cols`
bb_to_granges(api, ex_bed$id, extra_cols = c("column_name" = "character"))
```

`bb_to_granges` automatically assigns the column names and types for broad peak
and narrow peak files.

```{r bb_to_granges_narrowpeak, message=FALSE}
bed_id <- "bbad85f21962bb8d972444f7f9a3a932"
md <- bb_metadata(api, bed_id)
head(md)
bb_to_granges(api, bed_id)
```

`bb_to_granges` can also import big BED files.

```{r bb_to_granges_big_bed}
ex_bed <- bb_example(api, "bed")
bb_to_granges(api, ex_bed$id, "bigbed")
```

# Import a BEDset into a GRangesList

Create a GRangesList given a BEDset record identifier with `bb_to_grangeslist`.

```{r bb_to_grangeslist, message=FALSE}
bedset_id <- "lola_hg38_ucsc_features"
bb_to_grangeslist(api, bedset_id)
```

# Change genomic coordinate system with liftOver

Use `liftOver` to convert one genomic coordinate system to another.
Download the chain file from [UCSC](https://hgdownload.cse.ucsc.edu/downloads.html#liftover).

```{r liftOver, eval=FALSE}
if (!"BiocManager" %in% rownames(installed.packages())) {
    install.packages("BiocManager")
}
BiocManager::install("liftOver", dependencies = TRUE)
```

```{r convertCoordinates, message=FALSE}
id <- "7816f807ffe1022f438e1f5b094acf1a"
md <- bb_metadata(api, id, TRUE)
gro <- bb_to_granges(api, id)
gro

chain_url <- "https://hgdownload.cse.ucsc.edu/goldenPath/mm10/liftOver/mm10ToMm39.over.chain.gz"
tmpdir <- tempdir()
gz <- file.path(tmpdir, "mm10ToMm39.over.chain.gz")
download.file(chain_url, gz)
gunzip(gz, remove = FALSE)

ch <- import.chain(file.path(tmpdir, "mm10ToMm39.over.chain"))
seqlevelsStyle(gro) <- "UCSC"
gro39 <- liftOver(gro, ch)
gro39 <- unlist(gro39)
genome(gro39) <- "mm39"
gro39
```

# SessionInfo()

```{r sessionInfo}
sessionInfo()
```
