---
title: bedbaser
author:
- name: A Wokaty
  affiliation: CUNY School of Public Health
  email: jennifer.wokaty@sph.cuny.edu
package: bedbaser
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{bedbaser}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

# Introduction

```{r library, include=FALSE}
library(rtracklayer)
library(R.utils)
library(liftOver)
```

`r Biocpkg("bedbaser")` is an R API client for [BEDbase](https://bedbase.org)
that provides access to the [bedhost API](https://api.bedbase.org) and includes
convenience functions, such as to create GRanges and GRangesList objects.

# Install bedbaser and create a BEDbase instance

Install `r Biocpkg("bedbaser")` using `r CRANpkg("BiocManager")`.

```{r install, eval=FALSE}
if (!"BiocManager" %in% rownames(installed.packages())) {
    install.packages("BiocManager")
}
BiocManager::install("bedbaser")
```

```{r set-bedbaser-test_request, echo=FALSE, include=FALSE}
# We set .is_test_request to TRUE if !interactive() so we don't inflate
# statistics for BEDbase

options(bedbaser.test_request = !interactive())
```

Load the package and create a BEDbase instance, optionally setting the cache
to `cache_path`. If `cache_path` is not set, `r Biocpkg("bedbaser")` will
choose the default location.

```{r bedbase}
library(bedbaser)

bedbase <- BEDbase(tempdir())
```
`r Biocpkg("bedbaser")` can use the same cache as
[geniml](https://docs.bedbase.org/geniml/)'s BBClient by setting the
`cache_path` to the same location. It will create the following structure:

```
cache_path
    bedfiles
        a/f/afile.bed.gz
    bedsets
        a/s/aset.txt
```

# Convenience Functions

`r Biocpkg("bedbaser")` includes convenience functions prefixed with *bb_* to
facilitate finding BED files, exploring their metadata, downloading files, and
creating `GRanges` objects.

# BEDbase statistics

Use `bbs_stats()` to display the total available BED files, BEDsets, and
genomes. Set `detailed` to `TRUE` to display the type of BED formats and genomes
available.

## Find a BED file or BEDset

Use `bb_list_beds()` and `bb_list_bedsets()` to browse available resources in
BEDbase. Both functions display the id and names of BED files and BEDsets. An
id can be used to access a specific resource.

```{r bb_list_beds}
bb_list_beds(bedbase)

bb_list_bedsets(bedbase)
```

## Examine metadata

Use `bb_metadata()` to learn more about a BED or BEDset associated with an id.

```{r bb_metadata}
ex_bed <- bb_example(bedbase, "bed")
md <- bb_metadata(bedbase, ex_bed$id)
head(md)
```

## Show BED files in BEDset

Use `bb_beds_in_bedset()` to display the id of BEDs in a BEDset.

```{r bb_beds_in_bedset}
bb_beds_in_bedset(bedbase, "excluderanges")
```

## Search for a BED file by keyword

Search for BED files by keywords. `bb_bed_text_search()` returns all BED files
scored against a keyword query.

```{r bb_search}
bb_bed_text_search(bedbase, "cancer", limit = 10)
```

## Import a BED into a GRanges object

Create a GRanges object with a BED id with `bb_to_granges`, which
downloads and imports a BED file using `r Biocpkg("rtracklayer")`.

```{r bb_to_granges}
ex_bed <- bb_example(bedbase, "bed")
# Allow bedbaser to assign column names and types
bb_to_granges(bedbase, ex_bed$id, quietly = FALSE)
```

For BEDX+Y formats, a named list with column types may be passed through
`extra_cols` if the column name and type are known. Otherwise, `bb_to_granges`
guesses the column types and assigns column names.

```{r bb_to_granges_manual, eval=FALSE}
# Manually assign column name and type using `extra_cols`
bb_to_granges(bedbase, ex_bed$id, extra_cols = c("column_name" = "character"))
```

`bb_to_granges` automatically assigns the column names and types for broad peak
and narrow peak files.

```{r bb_to_granges_narrowpeak, message=FALSE}
bed_id <- "bbad85f21962bb8d972444f7f9a3a932"
md <- bb_metadata(bedbase, bed_id)
head(md)
bb_to_granges(bedbase, bed_id)
```

## Import a BEDset into a GRangesList

Create a GRangesList given a BEDset id with `bb_to_grangeslist`.

```{r bb_to_grangeslist, message=FALSE}
bedset_id <- "lola_hg38_ucsc_features"
bb_to_grangeslist(bedbase, bedset_id)
```

## Save a BED file

Save BED files or BEDsets with `bb_save`:

```{r bb_save}
bb_save(bedbase, ex_bed$id, tempdir())
```

# Accessing BEDbase API endpoints

Because `r Biocpkg("bedbaser")` uses the `r Biocpkg("AnVIL")` Service class,
it's possible to access any endpoint of the
[BEDbase API](https://api.bedbase.org/v1/docs).

```{r operations}
show(bedbase)
```

For example, to access a BED file's stats, access the endpoint with `$` and use
`r CRANpkg("httr")` to get the result. `show` will display information about the
endpoint.

```{r service_class_example}
library(httr)
show(bedbase$get_bed_stats_v1_bed__bed_id__metadata_stats_get)
id <- "bbad85f21962bb8d972444f7f9a3a932"
rsp <- bedbase$get_bed_stats_v1_bed__bed_id__metadata_stats_get(id)
content(rsp)
```

# Example: Change genomic coordinate system with liftOver

Given a BED id, we can use `r Biocpkg("liftOver")` to convert one genomic
coordinate system to another.

Install `r Biocpkg("liftOver")` and `r Biocpkg("rtracklayer")` then load the
packages.

```{r liftOver, eval=FALSE}
if (!"BiocManager" %in% rownames(installed.packages())) {
    install.packages("BiocManager")
}
BiocManager::install(c("liftOver", "rtracklayer"))

library(liftOver)
library(rtracklayer)
```

Create a GRanges object from a
[mouse genome](https://bedbase.org/bed/f2a5b06011706376560514c3f39648ea).
Create a BEDbase Service instance. Use the instance to create a GRanges
object from the BEDbase `id`.

```{r convertCoordinates_createGRangesObject, message=FALSE}
id <- "f2a5b06011706376560514c3f39648ea"
bedbase <- BEDbase()
gro <- bb_to_granges(bedbase, id)
gro
```

Download the
[chain file](https://hgdownload.cse.ucsc.edu/downloads.html#liftover) from UCSC.

```{r convertCoordinates_getTheChain, message=FALSE}
chain_url <- paste0(
    "https://hgdownload.cse.ucsc.edu/goldenPath/mm10/liftOver/",
    "mm10ToMm39.over.chain.gz"
)
tmpdir <- tempdir()
gz <- file.path(tmpdir, "mm10ToMm39.over.chain.gz")
download.file(chain_url, gz)
gunzip(gz, remove = FALSE)
```

Import the chain, set the sequence levels style, and set the genome for the
GRanges object.

```{r convertCoordinates_convert, message=FALSE}
ch <- import.chain(file.path(tmpdir, "mm10ToMm39.over.chain"))
seqlevelsStyle(gro) <- "UCSC"
gro39 <- liftOver(gro, ch)
gro39 <- unlist(gro39)
genome(gro39) <- "mm39"
gro39
```

```{r unset-bedbaser-test_request, echo=FALSE, include=FALSE}
# We unset bedbaser.test_request to FALSE

options(bedbaser.test_request = FALSE)
```

# SessionInfo()

```{r sessionInfo}
sessionInfo()
```
