---
title: BEDBaseR
author:
- name: J Wokaty
  affiliation: 
  - CUNY School of Public Health
  email: jennifer.wokaty@sph.cuny.edu
- name: Levi Waldron, PhD
  affiliation: 
  - Graduate School of Public Health and Health Policy, City University of New
    York, New York, NY, U.S.A.
  email: levi.waldron@sph.cuny.edu
package: BEDbaseR
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{BEDbaseR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r library, include=FALSE}
library(BEDbaseR)
```

BEDbaseR is an R client for the BEDbase API available at api.bedbase.org.

See available BED files with

```{r listRecords}
records <- listRecords("bed")
```

Download a BED file by its record identifier

```{r downloadBedFile}
# "GSM6856752_S1_H3K27ac_peaks"
record_id <- records$records[[1]]$record_identifier
filepath <- downloadBedFile(record_id, tempdir())
```

Ideally, we should be able to import with `r Biocpkg("rtracklayer")` to create
a GRanges object; however, the BEDfiles are in varying formats:

```{r importBed, eval=FALSE}
> import.bed(filepath)
# Error in scan(file = file, what = what, sep = sep, quote = quote, dec = dec,  : 
#   scan() expected 'an integer', got '0.9887082'
```

First try to use rtracklayer to import the BED file; however, if the BED file is
not in the format of the 3 required columns and 9 optional columns, we could try
to put it in a data.frame then arrange it into a GRanges object:
```{r fileToGRanges}
# Currently only using the 3 required columns, need to include more columns
gro <- fileToGRanges(filepath)
gro
```

Note: We could also do
```{r, eval=FALSE}
importToGRanges(record_id)
```

To Do:

* Import/include more of the optional columns into the GRanges object (See https://learn.gencore.bio.nyu.edu/ngs-file-formats/bed-format/)
* How to include nonstandard columns?

```
# hg19_UCSC_telomere.bed.gz
chr1	0	10000	10001	*	585	1	N	10000	telomere	no
chr1	249240621	249250621	10001	*	2486	2302	N	10000	telomere	no
chr2	0	10000	10001	*	585	1	N	10000	telomere	no
chr2	243189373	243199373	10001	*	2440	2027	N	10000	telomere	no
chr3	0	10000	10001	*	585	1	N	10000	telomere	no
chr3	198012430	198022430	10001	*	2095	1733	N	10000	telomere	no
chr4	0	10000	10001	*	585	1	N	10000	telomere	no
chr4	191144276	191154276	10001	*	2043	1658	N	10000	telomere	no
```

Development: https://github.com/users/jwokaty

Project Board: https://github.com/users/jwokaty/projects/4/views/1

```{r sessionInfo}
sessionInfo()
```   
