---
title: bedbaser
author:
- name: J Wokaty
  affiliation: 
  - CUNY School of Public Health
  email: jennifer.wokaty@sph.cuny.edu
package: bedbaser
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{bedbaser}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r library, include=FALSE}
library(bedbaser)
```

bedbaser is an R client for the BEDbase API available at https://api.bedbase.org.

## Install

```{r install, eval=FALSE}
if (!"BiocManager" %in% rownames(installed.packages()))
    install.packages("BiocManager")
BiocManager::install("bedbaser", dependencies=TRUE)
```
## Find a BED file or BEDset

BED files and BEDsets can be accessed by their record identifiers.

`get_records()` shows the record identifiers and names of available BED files
or BEDsets.

```{r get_records}
bedfiles <- get_records("bed")
bedfiles

bedsets <- get_records("bedset")
bedsets
```

`get_records_in_set()` returns the record identifiers of BED files in a
BEDset. 

```{r get_records_in_bedset}
get_records_in_set("excluderanges")
```

Download a BED file by its record identifier:

```{r downloadBedFile}
# "GSM6856752_S1_H3K27ac_peaks"
record_id <- bedfiles[[1,1]]
filepath <- downloadBedFile(record_id, tempdir())
```

Use `r Biocpkg("rtracklayer")` to import the BED file; however, if the BED file
is not in the format of the 3 required columns and 9 optional columns, we could
try to put it in a data.frame then arrange it into a GRanges object:

```{r fileToGRanges}
# Currently only using the 3 required columns, need to include more columns
gro <- fileToGRanges(filepath)
gro
```

Note: We could also do
```{r, eval=FALSE}
importToGRanges(record_id)
```

To Do:

* Import/include more of the optional columns into the GRanges object (See https://learn.gencore.bio.nyu.edu/ngs-file-formats/bed-format/)
* How to include nonstandard columns?

```
# GSM6856752_S1_H3K27ac_peaks.bed.gz
chr10	135001	135250	pos1	485	244.4055	0.9887082	11.85219
chr10	135751	135950	pos2	365	141.9129	1.362891	16.0
chr10	324251	324400	pos3	83	6.463647	3.682691	16.0
chr10	329701	329950	pos4	167	37.84343	2.141733	14.67584
chr10	394101	394550	pos5	898	198.678	2.176283	16.0
chr10	395301	395800	pos6	498	31.84901	3.966825	16.0
chr10	396401	396650	pos7	130	10.77274	3.593054	16.0
chr10	581051	581150	pos8	46	5.308168	3.115348	14.61217
chr10	625301	626100	pos9	1230	249.7136	2.300312	16.0
chr10	659601	660500	pos10	953	42.77816	4.47753	16.0
chr10	664301	664500	pos11	146	9.617266	3.924198	16.0
chr10	664901	665000	pos12	45	4.309098	3.384467	15.65356
chr10	665401	665550	pos13	95	17.92264	2.406144	16.0
chr10	665701	665850	pos14	111	22.07533	2.330053	16.0
chr10	683601	683850	pos15	180	19.07812	3.238006	16.0
```
Or

```
# hg19_UCSC_telomere.bed.gz
chr1	0	10000	10001	*	585	1	N	10000	telomere	no
chr1	249240621	249250621	10001	*	2486	2302	N	10000	telomere	no
chr2	0	10000	10001	*	585	1	N	10000	telomere	no
chr2	243189373	243199373	10001	*	2440	2027	N	10000	telomere	no
chr3	0	10000	10001	*	585	1	N	10000	telomere	no
chr3	198012430	198022430	10001	*	2095	1733	N	10000	telomere	no
chr4	0	10000	10001	*	585	1	N	10000	telomere	no
chr4	191144276	191154276	10001	*	2043	1658	N	10000	telomere	no
```

Development: https://github.com/users/jwokaty

Project Board: https://github.com/users/jwokaty/projects/4/views/1

```{r sessionInfo}
sessionInfo()
```   
